---
title: "Chap 2"
output:
  html_document:
    df_print: paged
  author: Xuewan Zhao
  html_notebook: default
  pdf_document: default
---
# Basic operations.
Create a vector in R.
Print sentences and objects by function cat().
```{r}
x = c(1,2,3)
y = c(4,5,6)
cat("x = ", x, "\n")
cat("y = ", y)
```

Do operations among vectors.
```{r}
length(x)
cat("x + y = ", x+y)
```


We can use function ls() to list all objects in this workspace.
rm() can be used to remove objects you no longer need in this workspace.
```{r}
ls()
rm(x)
ls()
# Remove all objects.
rm(list = ls())
ls()
```

We can use function matrix() to create a matrix. And do some operations.
```{r}
m = matrix(c(1,2,3,4),2,2,byrow=TRUE)
sqrt(m)
m^2
```

We can create random variables, and use summary() to observe the information for this object.
```{r}
set.seed(1303)
x=rnorm(50)
summary(x)
y=x+rnorm(50,mean=50,sd=.1)
cat("\nMean of x is", mean(x),"\nvariance of x is", var(x),
    "\nstandard deviation is", sd(x),"\nsquare root of variance is",sqrt(var(x)), 
    "\ncorrelation of x and y is",cor(x,y))
```

# Graphics.
```{r}
plot(x,y,xlab="this is the x-axis",ylab="this is the y-axis",
main="Plot of X vs Y")
jpeg("Figure.pdf")
plot(x,y,col="green")
dev.off()

```

```{r}
# We can create a sequence by function sep()
x=seq(-pi,pi,length=50)
y = x
f=outer(x,y,function(x,y)cos(y)/(1+x^2))
contour(x,y,f) # To plot explicitly, we need first run this line.
contour(x,y,f,nlevels=45,add=T)
fa=(f-t(f))/2 # t() means transpose here.
contour(x,y,fa,nlevels =15)

image(x,y,fa) # Heat map.
persp(x,y,fa)
persp(x,y,fa,theta=30)
persp(x,y,fa,theta=30,phi=20)
persp(x,y,fa,theta=30,phi=70)
persp(x,y,fa,theta=30,phi=40)
```

# Indexing data.
```{r}
A=matrix(1:16,4,4)
prmatrix(A)
cat("\nA[2,3] is\n")
prmatrix(A[2,3])
cat("\nA[c(1,3),c(2,4)] is\n")
prmatrix(A[c(1,3),c(2,4)])
cat("\nA[1:2,] is\n")
prmatrix(A[1:2,])
cat("\nA[,1:2] is\n")
prmatrix(A[,1:2])
cat("\nA[-c(1,3) ,] is\n")
prmatrix(A[-c(1,3) ,])
cat("\ndim(A) is",dim(A))
```

# Loading data.
```{r}
cur = getwd()
Auto=read.table(paste(cur,"/Datasets/Auto.data",sep = ""))
fix(Auto)
Auto=read.table(paste(cur,"/Datasets/Auto.data",sep = ""),header=T,na.strings ="?")
fix(Auto)
Auto=read.csv(paste(cur,"/Datasets/Auto.csv",sep = ""),header=T,na.strings ="?")
fix(Auto)
dim(Auto)
Auto=na.omit(Auto)
dim(Auto)
names(Auto)
```

# Additional Graphical and Numerical Summaries
```{r}
plot(Auto$cylinders , Auto$mpg)
attach(Auto) # not suggested.
plot(cylinders , mpg)
cylinders =as.factor(cylinders ) # convert numerical vector into variable.
plot(cylinders , mpg)
plot(cylinders , mpg, col="red")
plot(cylinders , mpg, col="red", varwidth=T)
plot(cylinders , mpg, col="red", varwidth=T,horizontal =T)
plot(cylinders , mpg, col="red", varwidth=T, xlab="cylinders ",
ylab="MPG")
hist(mpg)
hist(mpg,col=2) # col = 2 equals to col = "red"
hist(mpg,col=2,breaks=15)
pairs(Auto)
pairs(~mpg + displacement + horsepower + weight +acceleration , Auto)
plot(horsepower ,mpg)
identify(horsepower ,mpg,name) # didn't see it work.
```
